---
title: "Lab06 - multiple comparisons"
output: pdf_document
---

## Goals

The goal in this lab is to practice finding familywise conidence intervals for multiple comparisons.

## Loading packages

Here are some packages with functionality you may need for this lab.  Run this code chunk now.

```{r}
library(readr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(dplyr)
library(gmodels)

options("pillar.sigfig" = 10) # print 10 significant digits in summarize output

longevity <- read_csv("http://www.evanlray.com/data/sleuth3/ex0501_longevity.csv") %>%
  mutate(
    Diet = factor(Diet, levels = c("NP", "N/N85", "N/R50", "R/R50", "lopro", "N/R40"))
  )

nrow(longevity)
```

#### Find the multiplier that would be used for Bonferroni 95% familywise intervals for 5 comparisons based on this data set.

```{r}
1 - 0.05/(2 * 5)
qt(0.995, df = 349 - 6)
```

#### Find the multiplier that would be used for Scheffe 95% familywise intervals for 5 comparisons based on this data set.

```{r}
sqrt((6-1) * qf(0.95, df1 = 6-1, df2 = 349 - 6))
```


#### Find familywise 95% Confidence Intervals

The code below finds individual 95% confidence intervals for the 5 differences in group means the researchers planned for this study.  For each, add calculations of appropriately adjusted Bonferroni and Scheffe familywise intervals.

```{r}
anova_fit <- lm(Lifetime ~ Diet, data = longevity)
```

#### (a) $H_0: \mu_2 = \mu_3$ vs $H_A: \mu_2 \neq \mu_3$.  Are the population mean lifetimes the same for the \textbf{N/N85} and \textbf{N/R50} groups?

```{r}
fit.contrast(anova_fit, "Diet", c(0, 1, -1, 0, 0, 0), conf.int = 0.95)

# Bonferroni intervals
-9.605955 - 2.59 * 1.187682
-9.605955 + 2.59 * 1.187682

# Scheffe intervals
-9.605955 - 3.35 * 1.187682
-9.605955 + 3.35 * 1.187682
```

Here's an easier way to find the Bonferroni interval.  To get a familywise 95% CI for five intervals, each confidence interval will have an individual confidence level of $(1 - 0.05/5) \times 100\%$, or 99%.  Therefore, the following code to get an individual 99% CI will give us the first of 5 intervals with a familywise confidence level of 95%:

```{r}
fit.contrast(anova_fit, "Diet", c(0, 1, -1, 0, 0, 0), conf.int = 0.99)
```

Note that this result agrees with our manual calculation above, up to rounding errors.

#### (b) $H_0: \mu_3 = \mu_4$ vs $H_A: \mu_3 \neq \mu_4$.  Are the population mean lifetimes the same for the \textbf{N/R50} and \textbf{R/R50} groups?

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, -1, 0, 0), conf.int = 0.95)

# Bonferroni intervals
-0.5885312 - 2.59 * 1.19355
-0.5885312 + 2.59 * 1.19355

# Scheffe intervals
-0.5885312 - 3.35 * 1.19355
-0.5885312 + 3.35 * 1.19355
```

Or, the easier way for the Bonferroni interval:

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, -1, 0, 0), conf.int = 0.99)
```

#### (c) $H_0: \mu_3 = \mu_6$ vs $H_A: \mu_3 \neq \mu_6$.  Are the population mean lifetimes the same for the \textbf{N/R50} and \textbf{N/R40} groups?

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, 0, 0, -1), conf.int = 0.95)

# Bonferroni intervals
-2.819484 - 2.59 * 1.171097
-2.819484 + 2.59 * 1.171097

# Scheffe intervals
-2.819484 - 3.35 * 1.171097
-2.819484 + 3.35 * 1.171097
```


Or, the easier way for the Bonferroni interval:

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, 0, 0, -1), conf.int = 0.99)
```

#### (d) $H_0: \mu_3 = \mu_5$ vs $H_A: \mu_3 \neq \mu_5$.  Are the population mean lifetimes the same for the \textbf{N/R50} and \textbf{N/R50} lopro groups?

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, 0, -1, 0), conf.int = 0.95)

# Bonferroni intervals
2.611469 - 2.59 * 1.19355
2.611469 + 2.59 * 1.19355

# Scheffe intervals
2.611469 - 3.35 * 1.19355
2.611469 + 3.35 * 1.19355
```

Or, the easier way for the Bonferroni interval:

```{r}
fit.contrast(anova_fit, "Diet", c(0, 0, 1, 0, -1, 0), conf.int = 0.99)
```

#### (e) $H_0: \mu_2 = \mu_1$ vs $H_A: \mu_2 \neq \mu_1$.  Are the population mean lifetimes the same for the \textbf{N/N85} and \textbf{NP} groups?

```{r}
fit.contrast(anova_fit, "Diet", c(-1, 1, 0, 0, 0, 0), conf.int = 0.95)

# Bonferroni intervals
5.289187 - 2.59 * 1.301006
5.289187 + 2.59 * 1.301006

# Scheffe intervals
5.289187 - 3.35 * 1.301006
5.289187 + 3.35 * 1.301006
```

Or, the easier way for the Bonferroni interval:

```{r}
fit.contrast(anova_fit, "Diet", c(-1, 1, 0, 0, 0, 0), conf.int = 0.99)
```

#### What is the interpretation of the individual 95% confidence interval obtained in part (a) from the fit.contrast function?

We are 95% confident that the difference in population mean lifetimes for the \textbf{N/N85} and \textbf{N/R50} groups is between -11.942 and -7.270.  For 95% of samples, an interval calculated in this way will contain the difference in population mean lifetimes for the \textbf{N/N85} and \textbf{N/R50} groups.

#### What is the interpretation of the Bonferroni intervals?  

We are 95% confident that the difference in population mean lifetimes for the \textbf{N/N85} and \textbf{N/R50} groups is between -12.68 and -6.53, for the \textbf{N/R50} and \textbf{R/R50} groups is between -3.68 and 2.50, for the \textbf{N/R50} and \textbf{N/R40} groups is between -5.85 and 0.21, for the \textbf{N/R50} and \textbf{N/R50} lopro groups is between -0.48 and 5.70, and for the \textbf{N/N85} and \textbf{NP} groups is between 1.92 and 8.66.  For 95% of samples, all 5 of the intervals constructed in this way will simultaneously contain the difference in means they are estimating.





